---
title: "Data Wrangling with Dplyr"
author: "Peralta santos"
date: "11/13/2019"
output: html_document
---

```{r}
# Install if needed
#install.packages("tidyverse")
#install.packages("readr")
#install.packages("here")

library(tidyverse) # Versatile package for data analysis
library(readr) # read csv files
library(here) #easier to deal with relative paths 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse    = TRUE,
  warning     = FALSE, # Whether to display warnings
  message     = FALSE, # Whether to display messages
  echo        = TRUE,  # In this case we want to see the code 
  eval        = TRUE   # Whether to evaluate the code and include its results
  )

set.seed(1)       # Set the seed is important for getting reproducible reports 
options(scipen=2) # set the decimals 
```

About the dataset 

sns is a dataset of telephone triage, with the date, the type of call and the number 

```{r}
#Load the data
sns <- read_csv(here::here("data", "sns.csv"))
dim(sns)
sns 
```

# Data Wrangling with Dplyr

Dplyr aims to provide a function for each basic verb of data manipulation:

+ filter() to select cases based on their values.
+ arrange() to reorder the cases.
+ select() and rename() to select variables based on their names.
+ mutate() and transmute() to add new variables that are functions of existing variables.
+ summarise() to condense multiple values to a single value.
+ sample_n() and sample_frac() to take random samples.

more infro here -> https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html 

Deal with dates
```{r}
# Transform to date with Libridate 
sns$date <- as.Date(sns$periodo, "%Y-%m-%d")
sns$year <- as.factor(format(sns$date,'%Y'))
# we will cover this later 
sns <- sns %>%
  mutate(month_name=months(date)
  )
sns$week <- lubridate::isoweek(sns$date)
```

create a new variable with mutate and recode
```{r}
sns <- sns %>%
  mutate(type_call = recode(
    indicador, 
     "Encaminhamentos p/ Serviço Urgência" = "Emergency Care",
     "Encaminhamentos p/ Autocuidados" = "Self-care",
     "Encaminhamentos p/ CIAV" = "Poison Control hot-line",
     "Encaminhamentos p/ CSP" = "Primary care",
     "Encaminhamentos p/ INEM" = "Prehospital Emergency Care", 
     "Encaminhamentos p/ Outros" = "Other"
    ),
     number_calls = valor_absoluto
  )
```

use select to trim our dataset 
```{r}
sns <- sns %>%
  select(date,
         year,
         month_name,
         week,
         type_call,
         number_calls
  )
```

use filter to play with data 
```{r}
sns1 <- sns %>%
  filter(type_call =="Emergency Care"
  )

sns1 #that looks nice

# more complex filtering 
sns2 <- sns %>%
  filter(type_call =="Emergency Care" & 
           number_calls >1000
  )

sns2 #that looks nice
```

use group_by and summarize to colapse data 

```{r}
# Create a Sum variable daily
sns3 <-sns %>% group_by(date) %>% 
            summarise(calls_sum=sum(number_calls)
                      )

# Other way of summarizing the data 
# Create a Sum variable daily
sns4 <-sns %>% group_by(month_name) %>% 
            summarise(calls_mean=mean(number_calls),
                      calls_sd=sd(number_calls),
                      calls_max=max(number_calls),
                      calls_min=min(number_calls)
                      )

# now lets summarize by week
sns5 <-sns %>% group_by(week) %>% 
            summarise(calls_mean=mean(number_calls),
                      calls_sd=sd(number_calls),
                      calls_max=max(number_calls),
                      calls_min=min(number_calls)
                      )

# PRO TIPS
# The naming of the dataframes is horrible, they should be short and explanatory
sns_sum_week <- sns5
```

save the table 
```{r}
#This is for later use
save(sns4, file = "table_desc1.RData")
```

# Modeling 

```{r}
# Create a time variable 
sns3<- sns3 %>% 
  mutate(time = 1:n()
         )

# Create a cut off for a policy change 
sns3$start <-  ifelse(sns3$date >= "2017-10-01", 1,0)
```

```{r}
# create a model in this case a glm with a link long and a poisson family 
# its count data, it should be fine

m1 <- glm(
  calls_sum ~ time + 
    start, 
  family="poisson",
  data = sns3)
```

now lets create a tidy model
```{r}
library(broom)

table_m1 <- tidy(m1, 
                 conf.int=TRUE, 
                 exponentiate = TRUE)

table_m1
```

This is a format we can work with 
```{r}
table_m1<- table_m1 %>%
  dplyr::select(term, 
         estimate, 
         conf.low, 
         conf.high,
         p.value)

table_m1<- table_m1 %>%
  mutate(term = recode(term, 
     "(Intercept)" = "Baseline weekly n. calls",
     "time" = "Time trend",
     "start" = "Effect after 2017")
  )


```


save the table 
```{r}
#This is for later use
save(table_m1, file = "table_m1.RData")
```



Now lets calculate the model predicted values 
```{r}
pred_m1<- augment(m1)

# We are still in the log scale we need to exponentiate 
pred_m1$fit <- exp(pred_m1$.fitted)
pred_m1$UCI <- exp(pred_m1$.fitted + (1.96 * pred_m1$.se.fit) )
pred_m1$LCI <- exp(pred_m1$.fitted - (1.96 * pred_m1$.se.fit) )

# Tip
#Try to create the above variables using Dplyr mutate 
```



# Plot the data 
```{r}


plot.fit <- ggplot() +    # we leave it blank to have more flexibility
      #Model  Prediction
      geom_line(data = pred_m1,                # This creates the oberved valued line
               aes(x=time,               # we need to assign a != Y axis 
                   y = fit),
               color="red",              #type of plot, in this case is a line color red and dotted 
               geom="line",
               lwd=1,
               linetype="dotted") +
  # this create the shadow of the CI
  geom_ribbon(data = pred_m1,            
              aes(x=time,
                  ymin=LCI-1000,       # Our model is so bad that we need to add uncertancy to the CI 
                  ymax=UCI+1000),      # Our model is so bad that we need to add uncertancy to the CI
              alpha=0.5,
              fill="grey") + 
    #Observed values
  geom_line(data = pred_m1,               # This creates the oberved valued line
               aes(x=time,               # we need to assign a != Y axis 
                   y = calls_sum),   
                   color="darkcyan",     # and color 
               geom="line",
               lwd=1
            ) +
  #Create a vertical line with intercep x=44
    geom_vline(xintercept= 305, 
             colour="orange") + 
  #Create a text with intercep x=44
  geom_text(aes(x=305, 
                label="Policy change",y=30000), 
            colour="black", 
            angle=90, 
            vjust = 1.2, 
            size = 3)  +
  #Lets change the titles 
  labs(title = "Number of calls to telephone triage system",
           subtitle = "Observed and predicted valued",
           y = "Number of Calls",
           x = "Time in days",
           caption = "Source: Portuguese Ministry of Health API | (horrible) poisson distribution model")  +
theme_minimal() +
theme(plot.title=element_text( face="bold", 
                               hjust=0.5,
                               lineheight=1.2),
      plot.subtitle=element_text(hjust=0.5),
      legend.position="top",
      legend.direction="horizontal"
      )

plot.fit
ggsave( here::here("reports", "plot_model.png"))
```


The computational environment that was used to generate this version is as follows:
```{r echo = FALSE}
devtools::session_info()
```

We want the script in the reports folder
```{r}
rmarkdown::render('analysis_script.Rmd', 
                  output_file='analysis_script_final.html', 
                  output_dir=here::here("reports"))
```

#### The end 
