---
title: "R Markdown Basics"
author: "A Peralta Santos"
date: "`r Sys.Date()`"
params:
output:
  html_document:
    code_download: TRUE
---

## Set the markdown setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse    = TRUE,
  warning     = FALSE, # Whether to display warnings
  message     = FALSE, # Whether to display messages
  echo        = FALSE, # prevent printing of the R code that generated the plot 
  eval        = TRUE   # Whether to evaluate the code and include its results
  )

set.seed(1)       # Set the seed is important for getting reproducible reports 
options(scipen=2) # set the decimals 
```


### Create a inline URL

A nice example [here](https://www.statpower.net/Content/310/R%20Stuff/SampleMarkdown.html) 

## Equations 

Use 2 dollar signs for an equation sepatared from text
$$ E(log(Y_t)) = \beta_0 + \beta_{1-11} \sum_{m=1}^{m=12} MONTH_t + \beta_{12} TIME + \beta_{13} INT_t + \beta_{14} TIME*INT_t + log(POP_t) + \epsilon_t$$
Use a single dolar sign for math along side the text $E(log(Y_t))$ 


## Inline coding 

```{r}
x <- 2+2
```


The result of X is `r x` 

## Text formating 

```{r}
# to get some help install remedy package 
devtools::install_github("ThinkR-open/remedy") # Install remedy 
```


#### Footnote

The START RAs are the best ^[This is a footnote]

I love systematic reviews ^[this is another footnote]

#### Insert and URL
[this is an URL](https://www.washington.edu/research/research-centers/start-center/)

Write in **Bold**

Write in *Italic*


#### Create lists 

No special order

+  item 1
+  item 2
+  item 3 


Special order

1. item 1
2. item 2
3. item 3
 
Lets try kniting this markdown 

Lets try to knit again but this time with echo = TRUE in the settings

### Advanced stuff 
We will knit this file to a diferent directory

```{r}
library(here) #easier to deal with relative paths 

rmarkdown::render('markdown_basics.Rmd', 
                  output_file='test.html', 
                  output_dir=here::here("reports"))
```


#### The end


