---
title: "R Markdown Basics"
author: "A Peralta Santos"
date: "`r Sys.Date()`"
params:
output:
  html_document:
    code_download: TRUE
---

## Set the markdown setup 
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, # Whether to display warnings
  message = FALSE, # Whether to display messages
  echo = FALSE, # prevent printing of the R code that generated the plot
  eval = TRUE # Whether to evaluate the code and include its results
  )

set.seed(1) # Set the seed is important for getting reproducible reports 
rm(list=ls()) # Clean the envorment
options(scipen=4) # set the decimals 
```


### Create a inline URL
A nice example [here](https://www.statpower.net/Content/310/R%20Stuff/SampleMarkdown.html) 

## Equations 
Use 2 dollar signs for an equation sepatared from text
$$ E(log(Y_t)) = \beta_0 + \beta_{1-11} \sum_{m=1}^{m=12} MONTH_t + \beta_{12} TIME + \beta_{13} INT_t + \beta_{14} TIME*INT_t + log(POP_t) + \epsilon_t$$
Use a single dolar sign for math along side the text $E(log(Y_t))$ 


## Inline coding 

```{r}
x <- 2+2
```


The result of X is `r x`

## Text formating 

```{r}
# to get some help install remedy package 
devtools::install_github("ThinkR-open/remedy") # Install remedy 
```


#### Footnote

^[This is a footnote]
^[this is another footnote]

#### Insert and URL
[this is an URL](https://www.washington.edu/research/research-centers/start-center/)

Write in **Bold**

Write in *Italic*


#### Create lists 

No special order

+  item 1
+  item 2
+  item 3 


Special order

1. item 1
2. item 2
3. item 3
 



